Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-04-08 docs/ACM/paper.tex
Working on Related Work.

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}

@@

>>>>

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%\citestyle{acmauthoryear}

@@

Mainstream compilers have only recently added support to OpenMP 4.0's task
parallelism.
An implementation of \textsf{clang} supporting tasks was released in March 14th,
2014\footnote{\url{https://clang-omp.github.io/}}.
A few weeks later, in April 22nd, 2014, such support was also announced in
\textsf{gcc} 4.9.
Thus, because the necessary infra-structure for the implementation of tasks is
relatively new, currently there are no tools, other than the \Taskminer,
that annotate programs automatically with these directives.
Nevertheless, there exists a long string of research aiming at the automatic
parallelization of programs.
In this section, we examine some elements in this list that relate to our
work.

Directive-based code annotation standards provide programmers with an easy-to-use
parallel programming model.
Task-based extensions to these standards have resulted in OpenMP 4.X,
StarSs \cite{bellens:sc:2006, perez:iccc:2008, planas:hpca:2009,
tejedor:hpdc:2011} and OmpSs \cite{bueno:icpp:2011, duran:ppl:2011}  which have
been extensively used to task-parallelize applications. Such
programming models come with tools that help programmers find the
best annotations for their code. For example,  Tareador
\cite{ayguade:wcae:2015} enables a programmer, by means of a graphical interface,
to annotate sequential code, thus allowing the identification of potential task
parallelization opportunities.
Contrary to Tareador,  this paper proposes an approach that enables the
automatic insertion of OpenMP task annotations to relevant fragments of a
sequential program.

The techniques that we use to map program regions to tasks is similar to
analyses previously used in the generation of code for data-flow machines.
Data-flow programming was originally proposed as a candidate to enable
task-based parallelism \cite{agrawal:ipdps:2010, chan:spaa:2007,
gupta:micro:2011}.  Agrawal et al. \cite{agrawal:ipdps:2010}  extended
data-flow programming  with task input/output specifications in  Cilk++.
Vandierendonck et al.  \cite{vandierendonck:hotpar:2011} further extended Cilk++
with dependency clauses to facilitate the design of complex
parallelization patterns. Both groups presented a unified scheduler based on
fork-join parallelism \cite{vandierendonck:pact:2011} that enabled the
execution of task-based applications. Other approaches have also used
data-flow graphs to exploit parallelism, like Data-Driven Tasks
\cite{tasirlar:icpp:2011}, where the programmer can use put/await clauses to
determine task arguments before execution.
Function-based task parallelism was proposed by Gupta {\em et al.}
\cite{gupta:micro:2011} to use function arguments  as a way to
specify task  dependencies.
Notice that such research efforts focused on giving to programmers the tools to
{\em construct} parallel programs.
The automatic annotation of ordinary programs with data-flow constructs was not
among their goals.

Many systems have been developed to extract task parallelism from sequential
programs (semi) automatically.
Examples include OSCAR
(Optimally SCheduled Advanced multiprocessoR), Multigrain Parallelizing
Compiler \cite{ishizaka:journal:2000, kasahara:iwlcpc:2000},  MAPS (MPSoC
Application Programming  Studio) \cite{castrillon:tii:2013, ceng:dac:2008}, and
DiscoPoP (Discovery of Potential Parallelism) \cite{discopop, li:jss:2016}.
Besides parallelization systems, tools like Paraver \cite{extrae, paraver},
Aftermath~\cite{drebes:hipeac:2014}, DAGvis \cite{huynh:wvpa:2015}, and TEMANEJO
\cite{brinkmann:parco:2011, brinkmann:journal:2013, temanejo} were  designed to
enable performance analysis and visualization of task-based programs.
Although these tools help the programmer in adapting code
to run in parallel, they are not fully automatic.
For instance, the work that is the closest of ours, in reach and effectiveness,
in our opinion, is the suite of techniques proposed by Ravishankar {\em et
al.}~\cite{Ravishankar14}.
The type of irregular loops that they handle is impressive; however,
the lack of the runtime support \`{a} la OpenMP 4.X still requires them to
modify code before parallelization.
In their words: ``{\em For all benchmarks and applications, all functions were 
inlined, and arrays of structures were converted to structures of arrays for use
with our prototype compiler.}"
